%basic cover letter template
\documentclass{letter}

\oddsidemargin = .2in
\evensidemargin = .2in
\textwidth = 6.3in
\topmargin = -.5in
\textheight = 9in

\newcommand {\qed}{\mbox{$\Box$}}
\renewcommand {\iff}{\Longleftrightarrow}
\newcommand {\R}{\mathbb{R}}
\newcommand {\N}{\mathbb{N}}
\newcommand {\Q}{\mathbb{Q}}
\newcommand {\Z}{\mathbb{Z}}

\newcommand {\sub}{\mbox{SB}}

\address{Molly L. Lewis\\
Department of Psychology\\
Stanford University\\
Jordan Hall\\
450 Serra Mall\\
Stanford, CA  94305-2130\\
\\
mll@stanford.edu}
\begin{document}

\begin{letter}

% This letter should contain two (100-word or shorter) summaries: a c oncise paragraph to the editor indicating the scientific grounds why the paper should be considered for a topical, interdisciplinary journal rather than for a single-discipline or archival journal; and a separate, 100-word summary of the paper's appeal to a popular (non-scientific) audience.

% The cover letter should state clearly what is included as the submission, including number of words in the text and number of display items (figures, tables, boxes) in the print version of the paper; number of additional words in the text (full Methods and Extended Data legends) and number of Extended Data figures and tables for the online-only version; any Supplementary Information (specifying number of items and format); number of supporting manuscripts.

\opening{Editorial Board\\ 
Cognition\\
~\\
Dear Editors,}

Please accept the manuscript ``The length of words reflects their conceptual complexity,'' which I am submitting with my coauthor Michael C. Frank. The manuscript reports original data revealing a regularity in natural languages that we believe has not been described previously: longer words tend to refer to more complex meanings. In a series of ten studies, we find this regularity both in speakers' inferences about novel words, as well as in the lexicons of 80 languages. Our data suggest that pragmatic pressures may constrain the arbitrary character of the lexicon.

This manuscript is not currently submitted elsewhere and will not be submitted elsewhere prior to editorial decision.

APA ethical standards on the collection of data from human participants were followed. My coauthor and I do not have any interests that might be interpreted as influencing the research.

I will be serving as the corresponding author for this manuscript. My coauthor and I have agreed to the byline order and to submission of the manuscript in this form. I have assumed responsibility for keeping my co-author informed of our progress through the editorial review process, the content of the reviews, and any revisions made. I understand that, if accepted for publication, a certification of authorship form will be required that all authors will sign.

Sincerely,\\
\\
\\
Molly L. Lewis\\
Department of Psychology\\
Stanford University\\

\end{letter}

\end{document}