<h2> The length of words reflects their conceptual complexity </h2>
<h3>Supplementary Information </h3>

***
***

**TABLE OF CONTENTS**<br/>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;**Study 1: [Geon mapping task](#1)** <br/>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;**Study 2: [Geon complexity norms](#2)**<br/>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;**Study 3: [Geon mapping task control (random syllables)](#3)**<br/>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;**Study 4: [Real object complexity norms](#4)** <br/> 
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;**Study 5: [Real object mappping task](#5)**<br/> 
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;**Study 6: [Real object mapping task control (random syllables)](#6)** <br/> 
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;**Study 7: [Real object production task](#7)**<br/> 
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;**Study 8: [Geon study time task](#8)**<br/> 
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;**Study 9: [Real object study time task](#9)**<br/> 
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;**Study 10: [English complexity norms](#10)**<br/>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;**Study 11: [Cross-linguistic analysis](#11)** <br/> 
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;**Study 12: [Simultaneous frequency task](#12)**<br/>
&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;&nbsp;**Study 13: [Sequential frequency task](#12)**<br/>
  
This document was created from an R Markdown file. The R Markdown file can be found here. All analyses and plots can be reproduced from the raw data with the code in this file.

***
***

```{r global_vars, include=F}
## SET GLOBAL VARIABLES
rm(list=ls())
processNorms = TRUE # process norms or load norms? 
doSlow = FALSE # do time-consuming pre-processing steps?
fs = 15 # set graphical params
ts = 15
rs = 6

## LOAD LIBRARIES AND FUNCTIONS
library('psych')
library('boot')
library("rmarkdown")
source("/Documents/GRADUATE_SCHOOL/Ranalysis/useful.R")

# stuff for rendering to html in R markdown (need rmarkdown library too.)
library('RCurl')
library('bitops')
options(rpubs.upload.method="internal")
options(RCurlOptions = list(verbose = F,capath=system.file("CurlSSL", "cacert.pem", package="RCurl", ssl.verifypeer=F)))

# bootstrap CIs on CI (http://dodonovs.com/R/002-r.htm)
boot.cor = function(x, y, n=5000, p=0.95, method="pearson"){ 
  w = length(x); x.r = x; y.r = y 
  sm = 1:w; cor.b = 1:n 
  for (k in 1:n){ 
    s = sample(sm, w, replace = T) 
    for (i in 1:w) 
      { 
      x.r[i] = x[s[i]] 
      y.r[i] = y[s[i]] 
      } 
    cor.b[k] = cor(x.r, y.r, use = "pairwise", method) 
    } 
  cor.b = sort(cor.b); a = round(n*(1-p)/2,0); b = round(n*(p+1)/2,0) 
  vec = c(cor.b[a+1], cor(x, y, use = "pairwise", method), cor.b[b-1]) 
  n.r = c("value"); n.c = c("lower_bound", "correlation", "upper_bound") 
  matrix(vec,1,3,dimnames = list(n.r,n.c)) 
  }

# partial correlation
bm.partial<-function(x,y,z) {
  round((cor(x,y, use="complete.obs")-
           cor(x,z, use="complete.obs")*
           cor(y,z, use="complete.obs"))/
                                     sqrt((1-cor(x,z, use="complete.obs")^2)*(1-cor(y,z, use="complete.obs")^2)),4)
  }

# effect size for forced choice tasks
d.fc <- function(d) {
  cond <- all(intersect(levels(d$langCondition),c("\"long\"" , "\"short\"")) == c("\"long\"" , "\"short\""))
  
  # http://www.meta-analysis.com/downloads/Meta-analysis%20Converting%20among%20effect%20sizes.pdf
  if (cond) {
    d<- d[d$langCondition == "\"long\"" | d$langCondition ==  "\"short\"",]
    d <- droplevels(d)
    
    #use odds ratio to calculate d
    ns = table(d$langCondition, d$responseValue)
    or = (ns[1]*ns[4])/(ns[2]*ns[3]) 
    cf = sqrt(3)/pi
    effect_size = log(or) * cf #calculate d
    
    # caluclate 95 CI
    se = sqrt((1/ns[1]) + (1/ns[2]) + (1/ns[3]) + (1/ns[4])) # calculate se of log odds ratio
    d_se = se * (3/(pi^2)) # caclulate se of 
    d_err = d_se*1.96
    
    cill = effect_size - d_err
    ciul = effect_size + d_err
    rt.Mratio = mean(d$rt.ratio, na.rm = TRUE)
    c.Mratio = mean(d$c.ratio, na.rm = TRUE)
    
    es <- data.frame(effect_size=effect_size,
                     cill = cill,
                     ciul = ciul,
                     rt.Mratio = rt.Mratio,
                     c.Mratio = c.Mratio)
    }
  return (es)
  }

# proportions and CIs for forced choice tasks
p.fc <- function(d, dv){ 
  # get proportions
  complex_proport_c = sum(d$responseValue==dv)/length(d$responseValue)
  
  # get bootstrapped 95% CIs
  if (!is.na(complex_proport_c)) { 
    # bootstrap across subjects proportion responses for each category
    b <- boot(d$responseValue, function(u,i) table(u[i])[dv]/length(u), R = 1000) 
    ci <- boot.ci(b, type =  "basic")  
    ciwl = ci$basic[4]
    ciwu = ci$basic[5]
    
    es <- data.frame(p_complex = complex_proport_c,
                     ciwl = ciwl,
                     ciul = ciwu,
                     n=length(d$workerid))
    
  } else {
    es <- data.frame(p_c = NA,
                     ciwl = NA,
                     ciul = NA,
                     n=NA)
    
  }
  return (es)
}

# set working directory
setwd('/Documents/GRADUATE_SCHOOL/Projects/ref_complex/Papers/RC/analysis/')
```

A blurb about AMT?

<a name="1"/>
<h3> Study 1: Geon mapping task</h3>

The task can be found <a href="http://langcog.stanford.edu/expts/MLL/refComplex/Experiment38/ref_complex_38.html" target="_blank"> here</a>. 

The short word items were: "bugorn," "ratum," "lopus," "wugnum," "torun," "gronan," "ralex," "vatrus." The long word items were:  "tupabugorn," "gaburatum," "fepolopus," "pakuwugnum," "mipatorun," "kibagronan," "tiburalex," "binivatrus."

Across all studies presented here, some participants completed more than one study. The results presented here include the data from all participants, but all reported results remain reliable when excluding participants who completed more than one study. Participants were counted as a repeat participant if they completed a study using the same stimuli (e.g., completed both Studies 1 and 2 with geons).

Plotted below is the effect size (bias to select complex alternative in long vs. short word condition) as a function of the complexity ratio between the two object alternatives. Each dot corresponds to an object condition. Conditions are labeled by the quintiles of the two alternatives. For example, the "1/5" condition corresponds to the condition in which one alternative was from the first quintile and the other was from the fifth quintile. In the left plot, complexity is operationalized as the explicit complexity norms (Study 2). On the right, complexity is operationalized in terms of study times (Study 8). Effect sizes were calculated using the log odds ratio. In this and all subsequent plots, errors bars reflect  95% confidence intervals.

```{r 1:geon_mapping, warning=F, include=F}
## get data into long form
if (doSlow){
  # read in data
  d <- read.csv("data/RefComplex38.results",sep="\t",header=TRUE)
  
  # melt
  md <- melt(d,id.vars=c("workerid"),measure.vars=c(names(d)[c(grepl("_",names(d)))]))
  md$trial <- matrix(lapply(str_split(md$variable,"_"),function(x) {x[2]}))
  md$variable <- as.character(matrix(lapply(str_split(md$variable,"_"),function(x) {x[1]})))
  md$variable <- matrix(lapply(str_split(md$variable,"Answer."),function(x) {x[2]}))
  
  md$variable <- as.factor(as.character(md$variable))
  md$trial <- as.factor(as.character(md$trial))
  md$value <- as.factor(as.character(md$value))
  md$workerid <- as.factor(as.character(md$workerid))
  
  md$seq <- with(md, ave(value, workerid, variable, trial, FUN = seq_along))
  dc <- dcast(workerid + seq + trial ~ variable, data = md, value.var = "value")
  dc$seq <- NULL
  
  # make everything factors
  dc$criticalComplicated <- as.factor(as.numeric(gsub(
    "[[:punct:]]", "", dc$criticalComplicated))) #strip punctuations
  dc$criticalSimple <- as.factor(as.numeric(gsub(
    "[[:punct:]]", "", dc$criticalSimple))) #strip punctuations
  dc$langCondition <- as.factor(dc$langCondition)
  dc$objCondition <- as.factor(dc$objCondition)
  dc$response <- as.factor(dc$response)
  dc$responseSide <- as.factor(dc$responseSide)
  dc$responseValue <- as.factor(dc$responseValue)
  dc$word <- as.factor(dc$word)
  
  # merge in norms
  # see below for processing of these norms from raw data
  cg_norms = read.csv("data/complicatedNormsGeons_BYITEM.csv")  # Study 2
  rg_norms = read.csv("data/rtNormsGeons_BYITEM.csv") # Study 8

  # complicated
  index <- match(dc$criticalSimple, cg_norms$obj)
  dc$criticalSimple_c.norms <- cg_norms$meanRating[index]
  index <- match(dc$criticalComplicated,cg_norms$obj)
  dc$criticalComplicated_c.norms <- cg_norms$meanRating[index]
  
  # rt
  index <- match(dc$criticalSimple, rg_norms$obj)
  dc$criticalSimple_rt.norms <- rg_norms$log.rt [index]
  index <- match(dc$criticalComplicated, rg_norms$obj)
  dc$criticalComplicated_rt.norms <- rg_norms$log.rt [index]
  
  dc$c.ratio = dc$criticalSimple_c.norms/dc$criticalComplicated_c.norms
  dc$rt.ratio = dc$criticalSimple_rt.norms/dc$criticalComplicated_rt.norms
  
  write.csv(dc, "data/RC38_long.csv")
  }

dc <- read.csv('data/RC38_long.csv')
```

```{r, warning = F, echo = F, fig.height = 5, fig.width = 10}
## plot complexity ratios as a function of condition
# get effect sizes
de <- ddply(dc, .(objCondition), function (d) {d.fc(d)})

# get obj conditions
de$cond1 <- as.factor(unlist(matrix(lapply(str_split(de$objCondition ,"-"),function(x) {x[1]}))))
de$cond2 <- as.factor(unlist(matrix(lapply(str_split(de$objCondition ,"-"),function(x) {x[2]}))))
de$cond1<- as.factor(gsub("[[:punct:]]", "", de$cond1))
de$cond2<- as.numeric(gsub("[[:punct:]]", "", de$cond2))
de$objCondition2 = paste(de$cond1, "/", de$cond2, sep = "")

#fs = 20
#rs = 9
# plot complexity norm ratio
geon_c_plot = ggplot(de, aes(y=effect_size, x=c.Mratio)) +
  geom_pointrange(aes(ymax = cill, ymin=ciul))+
  geom_hline(yintercept=0,lty=2) +
  stat_smooth(method="lm") +
  ylab("Effect Size") +
  xlab("Complexity Rating Ratio") + 
  ggtitle("Effect size vs. complexity ratio") +
  theme(text = element_text(size=fs))  +
  scale_y_continuous(limits = c(-.33, .66)) +
  scale_x_continuous(limits = c(.25, 1.29)) +
  geom_text(aes(c.Mratio + .05, effect_size, label=objCondition2)) +
  annotate("text", x=1.15, y=.5, color = "red", size = rs,
           label=paste("r=",round(cor(de$effect_size, de$c.Mratio), 2))) +
  theme(text = element_text(size=fs),
        plot.title=element_text(size=ts, face = "bold"),
        plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = 'black')) 

# rt norm ratio
geon_rt_plot = ggplot(de, aes(y=effect_size, x=rt.Mratio)) +
  geom_pointrange(aes(ymax = cill, ymin=ciul))+
  geom_hline(yintercept=0, lty=2) +
  ggtitle("Effect size vs. study time ratio") +
  stat_smooth(method="lm") +
  ylab("Effect Size") +
  xlab("Study Time Ratio") + 
  theme(text = element_text(size=fs)) +  
  scale_y_continuous(limits = c(-.33, .66)) +
  scale_x_continuous(limits = c(.949, 1.005)) +
  geom_text(aes(rt.Mratio + .0025, effect_size, label=objCondition2)) +
  annotate("text", x=.997, y=.5, color = "red", size = rs,
           label=paste("r=",round(cor(de$effect_size, de$rt.Mratio), 2))) +
  theme(text = element_text(size = fs),
        plot.title = element_text(size=ts, face = "bold"),
        plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), 
        panel.border = element_blank(),
        axis.line = element_line(color = 'black')) 

#png("figure/geons_plot.png", height = 6, width = 12, units = "in", res=500)
multiplot(geon_c_plot, geon_rt_plot, cols = 2)
#dev.off()
```

```{r, include = FALSE}
## stats: correlations between norms and effect sizes
cor.test(de$c.Mratio, de$effect_size)
cor.test(de$rt.Mratio, de$effect_size)
```

<a name="2"/>
<h3> Study 2: Geon complexity norms</h3> 

The task can be found <a href="http://langcog.stanford.edu/expts/MLL/refComplex/Experiment34/ref_complex_34.html" target="_blank"> here</a>.

The relationship between number of geons and complexity rating is plotted below. Each dot corresponds to an object item (8 per condition). The x-coordinates have been jittered to avoid over-plotting.

```{r 2:geon_complexity_norms, warning = F, include = F}
## read in data and pre-process
if (processNorms) {
  # read in data
  d <- read.csv("data/RefComplex34.results",sep="\t",header=TRUE)
  
  # melt
  md <- melt(d,id.vars=c("workerid"),measure.vars=names(d)[grepl("obj",names(d))])
  md$trial <- matrix(lapply(str_split(md$variable,"_"),function(x) {x[2]}))
  md$obj <- unlist(matrix(lapply(str_split(md$value,"j"),function(x) {x[2]})))
  md$obj <- unlist(matrix(lapply(str_split(md$obj,".p"),function(x) {x[1]})))
  md$complexityLevel <- unlist(matrix(lapply(str_split(md$obj,"-"),function(x) {x[1]})))
  md$objID <- unlist(matrix(lapply(str_split(md$obj,"-"),function(x) {x[2]})))
  md$value <- NULL; md$variable <- NULL
  md = md[(md$trial != 0 & md$trial != 1),] # remove ball and circuit
  
  # get rating info
  mdr <- melt(d,id.vars=c("workerid"),measure.vars=names(d)[grepl("rating",names(d))])
  mdr$trial <- matrix(lapply(str_split(mdr$variable,"_"),function(x) {x[2]}))
  mdr$variable <- NULL
  
  # merge together based on trial and workerid
  m = merge(md, mdr, by=c("workerid","trial"))
  m$value <- as.numeric(as.character(m$value))
  
  # get norms by objects
  ms_all <- aggregate(value ~ obj, data=m, mean)
  ms_all$cih <- aggregate(value ~ obj, data=m, ci.high)$value
  ms_all$cil <- aggregate(value ~ obj, data=m, ci.low)$value
  
  names(ms_all)[2] = "meanRating"
  ms_all <- ms_all[order(ms_all$meanRating),]

  # save complexity by item
  write.csv(ms_all, "data/complicatedNormsGeons_BYITEM.csv")
  } 

cg_norms = read.csv("data/complicatedNormsGeons_BYITEM.csv")
```

```{r, echo = F, fig.width=4, fig.height=4}
## plot complexity norm by number of geons
# remove quotes from norms
cg_norms$obj <- as.factor(as.numeric(gsub("[[:punct:]]", "", cg_norms$obj)))
cg_norms$obj_class = as.numeric(substr(cg_norms$obj, 1, 1))
cg_norms$obj_item = as.numeric(substr(cg_norms$obj, 2, 2))

# make object class numeric
cg_norms$obj_class = as.numeric(cg_norms$obj_class) 

# plot
ggplot(cg_norms, aes(y=meanRating, x=jitter(obj_class))) +
  geom_pointrange(aes(ymax = meanRating+cih,ymin=meanRating-cil)) + 
  geom_smooth(method = "lm", color="blue", formula = y ~ x) +
  annotate("text", x=4, y=.25, color = "red", size=rs,
           label=paste("r=",round(cor(cg_norms$obj_class, cg_norms$meanRating), 2))) +
  xlab("Object condition") +
  ylab("Complexity Rating") +
  ggtitle("complexity rating vs. # of geons") +
  theme(text = element_text(size=fs),
        plot.title=element_text(size=ts, face = "bold"),
        plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = 'black')) 
```

```{r, include = F}
## stats: correlation between complexity norms and num geons (BY ITEM)
cor.test(cg_norms$obj_class, cg_norms$meanRating)
```

<a name="3"/>
<h3> Study 3: Geon mapping task control</h3> 

The task can be found <a href="http://langcog.stanford.edu/expts/MLL/refComplex/Experiment40/ref_complex_40.html" target="_blank"> here</a>.

Plotted below is the proportion complex object selections as a function of the number of syllables in the target label. The dashed line reflects chance selection between the simple and complex alternatives. 

```{r 3:geon_random_syllables, include = F}
## read in data 
d <- read.csv("data/RefComplex40.results",sep="\t",header=TRUE)

## get in long form and make everything factors
# get trial info
md <- melt(d, id.vars=c("workerid"),measure.vars=c(names(d)[c(grepl("_",names(d)))]))
md$trial <- matrix(lapply(str_split(md$variable,"_"),function(x) {x[2]}))
md$variable <- as.character(matrix(lapply(str_split(md$variable,"_"),function(x) {x[1]})))
md$variable <- matrix(lapply(str_split(md$variable,"Answer."),function(x) {x[2]}))

md$variable <- as.factor(as.character(md$variable))
md$trial <- as.factor(as.character(md$trial))
md$value <- as.factor(as.character(md$value))
md$workerid <- as.factor(as.character(md$workerid))

md$seq <- with(md, ave(value, workerid,  variable, trial, FUN = seq_along))
dc = dcast(workerid  + seq + trial ~ variable, data = md, value.var = "value")
dc$seq <- NULL

# make everything factors
dc$criticalComplicated <- as.factor(dc$criticalComplicated)
dc$criticalSimple   <- as.factor(dc$criticalSimple)
dc$langCondition   <- as.factor(dc$langCondition)
dc$objCondition   <- as.factor(dc$objCondition)
dc$response   <- as.factor(dc$response)
dc$responseSide   <- as.factor(dc$responseSide)
dc$responseValue   <- as.factor(dc$responseValue)
dc$word <- as.factor(dc$word)

# re-label length to be numeric
dc$len <- 1
dc$len[dc$langCondition=='"three"'] <- 3
dc$len[dc$langCondition=='"five"'] <- 5
```

```{r, warning = F, message = F, echo = F, fig.width = 4, fig.height = 4}
## plot by length condition
# get props
ms = ddply(dc ,.(len), function (d, dv) {p.fc(d,"\"complex\"")})

# plot
qplot(len,p_complex, position=position_dodge(),
      data=ms,geom="line",ylab="Prop. selection complex object", 
      xlab="Number of syllables")  +
  geom_linerange(aes(ymin=ciwl,ymax=ciul), position=position_dodge(.9)) +
  geom_point(aes(len,p_complex), position=position_dodge(.9)) +
  geom_abline(intercept = .5, slope = 0, linetype = 2) + #
  ylim(0,1) +
  ggtitle("Complex selections vs. # of syllables") +
  theme(text = element_text(size=fs),
        plot.title=element_text(size=ts, face = "bold"),
        plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = 'black'))
```

```{r include = F}
## stats: predict selections with word length
summary(glm(responseValue ~ len, data=dc, family = "binomial"))
```

<a name="4"/>
<h3> Study 4: Real object complexity norms</h3> 

The task can be found <a href="http://langcog.stanford.edu/expts/MLL/refComplex/Experiment9/ref_complex_9.html" target="_blank"> here</a>.

Plotted below is the correlation between the two samples (_N_ = 60 each) of complexity norms. Each dot corresponds to an object (_n_ = 60). 
```{r 4:objects_complexity_norms, include = FALSE}
## get data into long form, compute correlations between 2 samples, figure out quintiles, and remove dupliate subjects
if (processNorms){
  ## Sample #1
  da <- read.csv("data/RefComplex9a.results",sep="\t",header=TRUE)
  da[,31:92] <- lapply(da[,31:92],as.character)
  da[,31:92] <- lapply(da[,31:92],as.numeric)
  
  # melt
  md <- melt(da,id.vars=c("workerid"),measure.vars=names(da)[grepl("rating",names(da))])
  names(md) <- c("workerid", "rating", "value")
  ms = aggregate(value ~ rating, md, mean)

  ## Sample #2
  db <- read.csv("data/RefComplex9b.results",sep="\t",header=TRUE)
  db[,30:92] <- lapply(db[,30:92],as.character)
  db[,30:92] <- lapply(db[,30:92],as.numeric)
  
  # melt
  mdb <- melt(db,id.vars=c("workerid"),measure.vars=names(db)[grepl("rating",names(db))])
  names(mdb) <- c("workerid", "rating", "value")
  msb = aggregate(value ~ rating, mdb, mean)
  
  # merge two samples together to get lists for experiments 
  all_ratings = rbind(md, mdb)
  all_ms = aggregate(value ~ rating, all_ratings, mean)
  all_ms$ratingNum <- matrix(sapply(str_split(matrix(sapply(str_split(all_ms$rating,"rating"),
                                                            function(x) {x[2]})),"_"),function(x){x[1]}))
  all_ms$ratingNum<- as.numeric(str_replace_all(as.character(all_ms$ratingNum),"\\\"",""))
  all_ms$cil = aggregate(value ~ rating, all_ratings, ci.low)$value
  all_ms$cih = aggregate(value ~ rating, all_ratings, ci.high)$value
  
  all_ms = all_ms[!is.na(all_ms$ratingNum),] # get rid of ball and motherboard
  
  # add back in ratings for each samples
  all_ms = merge(all_ms, ms, by="rating")
  all_ms = merge(all_ms, msb, by="rating")
  
  all_ms$rating <- NULL
  names(all_ms) = c("meanRating", "ratingNum", "cil", "cih" ,"rating_1","rating_2")
  all_ms <- all_ms[c(2,1,3:6)]
  
  # get quintiles
  q = quantile(all_ms$meanRating, seq(0,1, by=.2))
  one = all_ms[which(all_ms$meanRating<q[2]), "ratingNum"]
  two = all_ms[which(all_ms$meanRating>q[2] & all_ms$meanRating<q[3]), "ratingNum"]
  three = all_ms[which(all_ms$meanRating>q[3] & all_ms$meanRating<q[4]), "ratingNum"]
  four = all_ms[which(all_ms$meanRating>q[4] & all_ms$meanRating<q[5]), "ratingNum"]
  five = all_ms[which(all_ms$meanRating>q[5]), "ratingNum"]
  
  one # 13 15 19 20 28 29  3 44 46 54 57 59
  two # 10 17  2 22 34 37  4 49  5 55  6  9
  three # 12 50 7 8 48 16 1 39 40 56 24 60
  four # 26 18 11 47 42 30 23 31 51 58 41 45
  five # 14 21 25 27 32 33 35 36 38 43 52 53
  
  all_ms$quintile = ifelse(is.element(all_ms$ratingNum, one), 1, 
                           ifelse(is.element(all_ms$ratingNum, two), 2,
                                  ifelse(is.element(all_ms$ratingNum, three), 3,
                                         ifelse(is.element(all_ms$ratingNum, four), 4,
                                                ifelse(is.element(all_ms$ratingNum, five), 5,
                                                       "error")))))
  all_ms$quintile = as.numeric(all_ms$quintile)
  all_ms <- all_ms[order(all_ms$meanRating),]
  
  # there is one participant who was in both samples. Look at correlation between samples without this participant in Sample #2 
  # Sample #2 
  # melt
  mdb <- melt(db[db$workerid != 'A1BQEX75BE1AYE',]
              ,id.vars=c("workerid"),measure.vars=names(db)[grepl("rating",names(db))])
  names(mdb) <- c("workerid", "rating", "value")
  msb = aggregate(value ~ rating, mdb, mean)
  
  co_norms_unique_sample = merge(ms, msb, by="rating")
  co_norms_unique_sample$ratingNum <- as.numeric(matrix(sapply(str_split(matrix(sapply(str_split(
    co_norms_unique_sample$rating,"rating"), function(x) {x[2]})),"_"),function(x){x[1]})))
  co_norms_unique_sample = co_norms_unique_sample[!is.na(co_norms_unique_sample$ratingNum),] # get rid of ball and motherboard

  co_norms_unique_sample$rating <- NULL
  names(co_norms_unique_sample) = c( "rating_1", "rating_2", "ratingNum")
  print(paste('reliability, removing duplicate subj:', round(cor(co_norms_unique_sample$rating_1, co_norms_unique_sample$rating_2),2))) # nearly identical to original sample
  
  # write to csv (use write.table to can exclude headers, so matlab can read for figure)
  write.table(all_ms, file="data/complicatedNormsObjs_BYITEM-m.csv", 
            row.names=FALSE, col.names=FALSE, sep=",")
  
  # write to csv
  write.csv(all_ms, "data/complicatedNormsObjs_BYITEM.csv")
  }

co_norms = read.csv("data/complicatedNormsObjs_BYITEM.csv")
```

```{r echo = F, fig.width = 4, fig.height = 4, warning = FALSE, message = FALSE, output = FALSE}
## plot correlation between two samples
qplot(rating_1,rating_2, position=position_dodge(),
      data=co_norms, geom="point", ylab="Sample #2", xlab="Sample #1") +
  geom_point(aes(rating_1,rating_2), position=position_dodge(.9)) +
  stat_smooth(method="lm") +
  ylim(0,1) +
  xlim(0,1) +
  ggtitle("Complexity rating samples") +
  annotate("text", x=.75, y=.2, color = "red", size = rs,
           label=paste("r=", round(cor(co_norms$rating_1, co_norms$rating_2), 2))) +
  theme(text = element_text(size=fs),
          plot.title=element_text(size=ts, face = "bold"),
          plot.background = element_blank(),
          panel.grid.major = element_blank(),
          panel.grid.minor = element_blank(),
          panel.border = element_blank(),
          axis.line = element_line(color = 'black'))
```

```{r, include = F}
## stats: correlation between sample #1 and #2
cor.test(co_norms$rating_1, co_norms$rating_2)
```

<a name="5"/>
<h3> Study 5: Real object mapping task</h3> 

The task can be found <a href="http://langcog.stanford.edu/expts/MLL/refComplex/Experiment35/ref_complex_35.html" target="_blank"> here</a>.

The linguistic items were identical to Study 1.

Plotted below is the effect size (bias to select complex alternative in long vs. short word condition) as a function of the complexity ratio between the two object alternatives. Each dot corresponds to an object condition. In the left plot, complexity is operationalized as the explicit complexity norms (Study 4). In the right plot, complexity is operationalized in terms of study times (Study 9). 

```{r 5:objects_mapping, warning = F, include = F}
## read in data get into long form
if (doSlow) {
  # read in data 
  d <- read.csv("data/RefComplex35.results", sep="\t", header=TRUE)
  
  # get in long form
  # get trial info
  md <- melt(d,id.vars=c("workerid"),
             measure.vars=c(names(d)[c(grepl("_",names(d)))]))
  md$trial <- matrix(lapply(str_split(md$variable,"_"),function(x) {x[2]}))
  md$variable <- as.character(matrix(lapply(str_split(md$variable,"_"),function(x) {x[1]})))
  md$variable <- matrix(lapply(str_split(md$variable,"Answer."),function(x) {x[2]}))
  
  md$variable <- as.factor(as.character(md$variable))
  md$trial <- as.factor(as.character(md$trial))
  md$value <- as.factor(as.character(md$value))
  md$workerid <- as.factor(as.character(md$workerid))
  
  md$seq <- with(md, ave(value, workerid,  variable, trial, FUN = seq_along))
  dc = dcast(workerid + seq + trial ~ variable, data = md, value.var = "value")
  dc$seq <- NULL
  
  write.csv(dc, "data/RC35_long.csv")
  }

dc <- read.csv("data/RC35_long.csv")
```

```{r, include = F}
## make everything factors
dc$criticalComplicated= gsub(" ", "", gsub("[[:punct:]]", "", dc$criticalComplicated))
dc$criticalSimple= gsub(" ", "", gsub("[[:punct:]]", "", dc$criticalSimple))

dc$criticalComplicated <- as.factor(dc$criticalComplicated)
dc$criticalSimple <- as.factor(dc$criticalSimple)
dc$langCondition <- as.factor(dc$langCondition)
dc$objCondition <- as.factor(dc$objCondition)
dc$response <- as.factor(dc$response)
dc$responseSide <- as.factor(dc$responseSide)
dc$responseValue <- as.factor(dc$responseValue)
dc$word <- as.factor(dc$word)
dc$trial <- as.numeric(dc$trial)
```

```{r, include = F}
## merge in norms
index <- match(dc$criticalSimple, co_norms$ratingNum)
dc$criticalSimple_c.norms <- co_norms$meanRating[index]
index <- match(dc$criticalComplicated, co_norms$ratingNum)
dc$criticalComplicated_c.norms <- co_norms$meanRating[index]

rto_norms = read.csv("data/rtNormsObjs_BYITEM.csv") #analyzed from raw data below (Study 9)
index <- match(dc$criticalSimple, rto_norms$Answer.train_image)
dc$criticalSimple_rt.norms <- rto_norms$log.rt [index]
index <- match(dc$criticalComplicated, rto_norms$Answer.train_image)
dc$criticalComplicated_rt.norms <- rto_norms$log.rt [index]

dc$c.ratio = dc$criticalSimple_c.norms/dc$criticalComplicated_c.norms
dc$rt.ratio = dc$criticalSimple_rt.norms/dc$criticalComplicated_rt.norms
```

```{r, warning = F, echo = F, fig.height = 5, fig.width = 10}
## plot complexity ratios vs. condition
# get effect sizes
de <- ddply(dc, .(objCondition), function (d) {d.fc(d)})

# get obj conditions
de$cond1 <- as.factor(unlist(matrix(lapply(str_split(de$objCondition ,"-"),function(x) {x[1]}))))
de$cond2 <- as.factor(unlist(matrix(lapply(str_split(de$objCondition ,"-"),function(x) {x[2]}))))
de$cond1<- as.factor(gsub("[[:punct:]]", "", de$cond1))
de$cond2<- as.factor(gsub("[[:punct:]]", "", de$cond2))
de$objRatio = as.numeric(de$cond1)/as.numeric(de$cond2)
de$l.objRatio <- log(de$objRatio)

de$objCondition2 = paste(de$cond1, "/", de$cond2, sep = "")

#fs = 20
#rs = 9 
# complexity plot
obj_c_plot = ggplot(de, aes(y=effect_size, x=c.Mratio)) +
  geom_pointrange(aes(ymax = cill, ymin=ciul)) +
  geom_hline(yintercept=0,lty=2) +
  stat_smooth(method="lm") +
  ylab("Effect Size") +
  xlab("Complexity Rating Ratio") + 
  scale_y_continuous(limits = c(-.33, .66)) +
  geom_text(aes(c.Mratio + .04, effect_size, label=objCondition2)) +
  annotate("text", x=1, y=.5, col = "red",
           label=paste("r=",round(cor(de$effect_size, de$c.Mratio, use = "complete"), 2)),
           size = rs) +
  ggtitle("Effect size vs. complexity ratio") + 
  theme(text = element_text(size=fs), 
        plot.title = element_text(size=ts, face = "bold"),
        plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = 'black')) 

# RT ratio plot
obj_rt_plot = ggplot(de, aes(y=effect_size, x=rt.Mratio)) +
  geom_pointrange(aes(ymax = cill, ymin=ciul)) +
  geom_hline(yintercept=0,lty=2) +
  stat_smooth(method="lm") +
  ylab("Effect Size") +
  xlab("Study Time Ratio") +
  geom_text(aes(rt.Mratio + .001, effect_size, label=objCondition2)) +
  annotate("text", x=1, y=.5, col = "red",
           label=paste("r=",round(cor(de$effect_size, de$rt.Mratio, use = "complete"), 2)),
           size = rs) +  
  ggtitle("Effect size vs. study time ratio") +
  theme(text = element_text(size=fs), 
        plot.title = element_text(size=ts, face = "bold"),
        plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = 'black')) +
        scale_y_continuous(limits = c(-.33, .66)) 
  
#png("figure/obj_plot.png", height = 6, width = 12, units = "in", res=500)
multiplot(obj_c_plot, obj_rt_plot, cols = 2)
#dev.off()

```

<a name="6"/>
<h3> Study 6: Real object mapping task control</h3> 

The task can be found <a href="http://langcog.stanford.edu/expts/MLL/refComplex/Experiment41/ref_complex_41.html" target="_blank"> here</a>.

Plotted below is the proportion of complex object selections as function of number of syllables. The dashed line reflects chance selection between simple and complex alternatives.

```{r 6:objects_random_syllables, warning = F, message = F, output = F, include = F}
## read data, and pre-process
d <- read.csv("data/RefComplex41.results",sep="\t",header=TRUE)

# get trial info
md <- melt(d,id.vars=c("workerid"),measure.vars=c(names(d)[c(grepl("_",names(d)))]))
md$trial <- matrix(lapply(str_split(md$variable,"_"),function(x) {x[2]}))
md$variable <- as.character(matrix(lapply(str_split(md$variable,"_"),function(x) {x[1]})))
md$variable <- matrix(lapply(str_split(md$variable,"Answer."),function(x) {x[2]}))

md$variable <- as.factor(as.character(md$variable))
md$trial <- as.factor(as.character(md$trial))
md$value <- as.factor(as.character(md$value))
md$workerid <- as.factor(as.character(md$workerid))

# get into long form
md$seq <- with(md, ave(value, workerid,  variable, trial, FUN = seq_along))
dc = dcast(workerid  + seq + trial ~ variable, data = md, value.var = "value")
dc$seq <- NULL

# make everything factors
dc$criticalComplicated <- as.factor(dc$criticalComplicated)
dc$criticalSimple   <- as.factor(dc$criticalSimple)
dc$langCondition   <- as.factor(dc$langCondition)
dc$objCondition   <- as.factor(dc$objCondition)
dc$response   <- as.factor(dc$response)
dc$responseSide   <- as.factor(dc$responseSide)
dc$responseValue   <- as.factor(dc$responseValue)
dc$word <- as.factor(dc$word)

# re-label length to be numeric
dc$len <- 1
dc$len[dc$langCondition=='"three"'] <- 3
dc$len[dc$langCondition=='"five"'] <- 5
```

```{r, warning = F, message = F, output = F, echo=F, fig.width = 4, fig.height = 4}
## plot proportion complex selections by length condition
# get props
ms = ddply(dc ,.(len), function (d, dv) {p.fc(d,"\"complex\"")})

#plot
qplot(len,p_complex, position=position_dodge(),
      data=ms,geom="line",ylab="Prop. selection complex object", 
      xlab="Number of syllables")  +
  geom_linerange(aes(ymin=ciwl,ymax=ciul), position=position_dodge(.9)) +
  geom_point(aes(len,p_complex), position=position_dodge(.9)) +
  geom_abline(intercept = .5, slope = 0, linetype = 2) + #
  ylim(0,1) +
  ggtitle("Complex selections vs. # of syllables") +
  theme(text = element_text(size=fs),
        plot.title=element_text(size=ts, face = "bold"),
        plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = 'black'))
```

```{r, include = F}
## stats: predict reference selection by word length
summary(glm(responseValue ~ len, data=dc, family = "binomial"))
```

<a name="7"/>
<h3> Study 7: Real object production task</h3> 

The task can be found <a href="http://langcog.stanford.edu/expts/MLL/refComplex/Experiment27/ref_complex_27.html" target="_blank"> here</a>.

There were 26 productions (4%) that included more than one word. These productions were excluded. 

For each object, we analyzed the log length of the production in characters as a function of the complexity norms (Study 4, left below). Length of production was correlated with the complexity norms: Longer labels were coined for objects that were rated as more complex (_r_=.17, _p_<.0001). 

We also analyzed the log length of the production (in characters) as a function of study times (Study 9, right below). Length of production was correlated with study times: longer labels were cointed for objects that were studied longer (_r_=.16, _p_<.001). 

```{r 7:objects_production_labels, warning = F, message = F, output = F, include = F}
## read in data and get into long form
d <- read.csv("data/RefComplex27.results", sep="\t", header=TRUE)
  
n <- names(d)
d$Answer.pic_1 = as.factor(as.character(d$Answer.pic_1))
d$Answer.pic_2 = as.factor(as.character(d$Answer.pic_2))
d$Answer.pic_3 = as.factor(as.character(d$Answer.pic_3))
d$Answer.pic_4 = as.factor(as.character(d$Answer.pic_4))
d$Answer.pic_5 = as.factor(as.character(d$Answer.pic_5))
d$Answer.pic_6 = as.factor(as.character(d$Answer.pic_6))
d$Answer.pic_7 = as.factor(as.character(d$Answer.pic_7))
d$Answer.pic_8 = as.factor(as.character(d$Answer.pic_8))
d$Answer.pic_9 = as.factor(as.character(d$Answer.pic_9))
d$Answer.pic_10 = as.factor(as.character(d$Answer.pic_10))

cols = c(n[grepl("cond",n)], n[grepl("pic",n)], n[grepl("descLength",n)] )
md1 <- melt(d,id.vars=c("workerid"), measure.vars=cols,na.rm=TRUE)
md1$trial <- as.numeric(matrix(lapply(str_split(md1$variable,"_"),function(x) {x[2]})))

md1 <- melt(d,id.vars=c("workerid"), measure.vars=n[grepl("pic",n)],na.rm=TRUE)
md1$trial <- as.numeric(matrix(lapply(str_split(md1$variable,"_"),function(x) {x[2]})))
names(md1)[3] = "picture"
md1 = md1[,-2]

md2 <- melt(d,id.vars=c("workerid"), measure.vars=n[grepl("desc_",n)],na.rm=TRUE)
md2$trial <- as.numeric(matrix(lapply(str_split(md2$variable,"_"),function(x) {x[2]})))
names(md2)[3] = "description"
md2 = md2[,-2]

md3 <- melt(d,id.vars=c("workerid"), measure.vars=n[grepl("descLength",n)],na.rm=TRUE)
md3$trial <- as.numeric(matrix(lapply(str_split(md3$variable,"_"),function(x) {x[2]})))
names(md3)[3] = "length"
md3 = md3[,-2]

md4 <- melt(d,id.vars=c("workerid"), measure.vars=n[grepl("cond",n)],na.rm=TRUE)
md4$trial <- as.numeric(matrix(lapply(str_split(md4$variable,"_"),function(x) {x[2]})))
names(md4)[3] = "condition"
md4 = md4[,-2]

md12 <- join(md1, md2,type = "inner")
md123 <- join(md12, md3,type = "inner")
md <- join(md123, md4,type = "inner")

# add columns
md$numWords = sapply(gregexpr("\\W+", md$description), length) - 1
md <- md[md$numWords == 1,] #remove multi word responses
md$log.length <- log(md$length) 
md$log.trial <- log(md$trial)
```

```{r, include = F}
## look at relationship between condition and label length
# aggregate across participants
md.agg <- ddply(md, .(workerid,condition), function (d) {mean(d$log.length)})
md.agg$condition = as.factor(md.agg$condition)
names(md.agg)[3] = "log.length"
md.agg = md.agg[md.agg$workerid != "A24IF41IP6LT55" ,] # one participant doesn't have both simple and complex (due to multi word responses)

## stats: paired t-test of length between simple and complex conditions 
t.test(md.agg[md.agg$condition == '"complex"',"log.length"],
       md.agg[md.agg$condition == '"simple"',"log.length"], paired = TRUE)
summary(lmer(log.length~condition + (1+log.trial|workerid), md))
```

```{r, include = F}
## look at relationship between complicated norms and label length
# merge and aggregate
index <- match(md$picture, co_norms$ratingNum)
md$c.norms <- co_norms$meanRating[index]

ms <- aggregate(log.length ~ c.norms + picture, data=md, mean)
ms$cih <- aggregate(log.length ~ c.norms + picture, data=md, ci.high)$log.length
ms$cil <- aggregate(log.length ~ c.norms + picture, data=md, ci.low)$log.length

# plot
obj_c_p_plot = ggplot(ms, aes(c.norms,log.length)) +
  geom_point() + 
  geom_smooth(method = "lm", color="blue", formula = y ~ x) +
  geom_errorbar(aes(ymax=log.length+cih,ymin=log.length-cil), size=0.2) +
  xlab("Object Complexity Norms") +
  ylab("Log Word Length (characters)") +
  annotate("text", x=.755, y=1.7, color = "red", size = rs,
    label=paste("r=",round(cor(md$log.length,md$c.norms), 2))) +
  ggtitle('Label length vs. Complicated norms') +
  theme(text = element_text(size=fs),
        plot.title=element_text(size=ts, face = "bold"),
        plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = 'black'))
```

```{r, include = F}
## stats: relationship between length and complexity norms
cor.test(md$log.length, md$c.norms)
summary(lmer(scale(log.length)~scale(c.norms) + (1+trial|workerid), md))
```

```{r, echo = F, fig.width = 8, fig.height = 4}
## look at relationship between study time and label length
# merge and aggregate
index <- match(md$picture, rto_norms$Answer.train_image)
md$rt.norms <- rto_norms$log.rt[index]

ms <- aggregate(log.length ~ rt.norms + picture, data=md, mean)
ms$cih <- aggregate(log.length ~ rt.norms + picture, data=md, ci.high)$log.length
ms$cil <- aggregate(log.length ~ rt.norms + picture, data=md, ci.low)$log.length

# plot
obj_rt_p_plot = ggplot(ms, aes(rt.norms,log.length)) +
  geom_point() + 
  geom_smooth(method = "lm", color="blue", formula = y ~ x) +
  geom_errorbar(aes(ymax=log.length+cih,ymin=log.length-cil), size=0.2) +
  xlab("Object Study Time") +
  ylab("Log Word Length (characters)") +
  annotate("text", x=7.5, y=1.7, color = "red", size = rs,
    label=paste("r=",round(cor(md$log.length,md$rt.norms), 2))) +
  ggtitle('Label length vs. Study time') +
  theme(text = element_text(size=fs),
        plot.title=element_text(size=ts, face = "bold"),
        plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = 'black'))
  
multiplot(obj_c_p_plot, obj_rt_p_plot, cols = 2)
```

```{r, include = F}
## stats: relationship between length and RT norms
cor.test(md$log.length, md$rt.norms)
summary(lmer(scale(log.length)~scale(rt.norms) + (1+trial|workerid), md))
```

<a name="8"/>
<h3> Study 8: Geon study time task</h3> 

The task can be found <a href="http://langcog.stanford.edu/expts/MLL/refComplex/Experiment37/ref_complex_37.html" target="_blank"> here</a>.

We excluded subjects who performed below chance on the memory task (fewer than 20 correct out of 40). A response was counted as correct if it was a correct rejection or a hit. This excluded 9 subjects (4%). With these participants excluded, the mean correct was 72%. 

Participants were also excluded based on study times. We transformed the time into log space, and excluded responses that were 2 standard deviations above or below the mean. This excluded 4% of responses. A histogram of study times is presented below. The solid line indicates the mean, and the dashed lines indicate two standard deviations above and below the mean.

```{r 8:geon_rt_norms, warning = F, echo = F, fig.width = 4, fig.height = 4, message = F}
## get data into long form (accuracy and RT df) and do exclusions
#if (processNorms) {
  # read in data
  raw <- read.csv("data/RefComplex37.results",sep="\t",header=TRUE)
  
  # look at accuracy, and exclude those below chance
  #boxplot(raw$Answer.correct, main = "distribution before accuracy exclusions")
  raw = raw[raw$Answer.correct > 20,]

  # prep accuracy and RT dataframes
  # accuracy data frame
  # melt
  n <- names(raw)
  cols = c( n[grepl("test",n)])
  mda <- melt(raw,id.vars=c("workerid"), measure.vars=cols,na.rm=TRUE)
  mda$trial <-as.numeric(matrix(lapply(str_split(mda$variable,"_"),function(x) {x[3]})))
  mda$var1 <- matrix(lapply(str_split(mda$variable,"_"),function(x) {x[1]}))
  mda$var2 <- matrix(lapply(str_split(mda$variable,"_"),function(x) {x[2]}))
  mda$var <- paste(mda$var1, mda$var2, sep = "_")
  mda$variable <- NULL; mda$var1 <- NULL; mda$var2 <- NULL
  
  mda$seq <- with(mda, ave(value, workerid, var, FUN = seq_along))
  da = dcast(workerid + seq + trial ~ var, data = mda, value.var = "value")
  da$seq <- NULL
  
  da=da[!is.na(da$trial),]
  
  da$Answer.test_answer = as.factor(da$Answer.test_answer)
  da$Answer.test_answerEval = as.factor(da$Answer.test_answerEval)
  da$Answer.test_image  = as.factor(da$Answer.test_image)
  
  # look at memory performance (correct: CR or hit)
  numCR = length(which(da$Answer.test_answerEval == "\"CR\""))
  numH = length(which(da$Answer.test_answerEval == "\"H\""))

  correct = (numCR + numH)/ dim(da)[1]
  #print(paste("percent correct:", round(correct,2)))
  
  # get obj condition variable
  da$test_image2 <- as.character(matrix(lapply(str_split(da$Answer.test_image,"j"), function(x) {x[2]})))
  da$objCondition <- as.factor(as.character(matrix(lapply(str_split(da$test_image2,"-"), function(x) {x[1]}))))
  da$objItem <- as.character(matrix(lapply(str_split(da$test_image2,"-"), function(x) {x[2]})))
  da$objItem<- as.factor(as.numeric(gsub("[[:punct:]]", "", da$objItem)))
  da$test_image2 <- NULL
  
  # look at memory performance (correct: CR or hit)
  numCR = length(which(da$Answer.test_answerEval == "\"CR\""))
  numH = length(which(da$Answer.test_answerEval == "\"H\""))

  correct = (numCR + numH)/ dim(da)[1]
  #print(paste("percent correct:" , round(correct,2)))
  
  # RT dataframe
  # melt
  n <- names(raw)
  cols = c(n[grepl("train", n)])
  cols = cols[1:40]
  md <- melt(raw, id.vars=c("workerid"), measure.vars=cols)
  md$trial <- as.numeric(matrix(lapply(str_split(md$variable,"_"),function(x) {x[3]})))
  md$var1 <- matrix(lapply(str_split(md$variable,"_"),function(x) {x[1]}))
  md$var2 <- matrix(lapply(str_split(md$variable,"_"),function(x) {x[2]}))
  md$var <- paste(md$var1, md$var2, sep = "_")
  md$variable <- NULL; md$var1 <- NULL; md$var2 <- NULL
  
  md$seq <- with(md, ave(value, workerid, var, FUN = seq_along))
  d = dcast(workerid + seq + trial ~ var, data = md, value.var = "value")
  d$Answer.train_NA <- NULL; d$seq <- NULL
  
  d=d[!is.na(d$trial),]
  d=d[!is.na(d$Answer.train_image),]
  d=d[!is.na(d$Answer.train_rt),]
  d=d[d$Answer.train_image != "undefined",]
  d=d[d$Answer.train_image != "",]
  
  d$Answer.train_rt = as.numeric(d$Answer.train_rt)
  
  # exclude outliers 2 standard deviations above and below mean (in log space)
  d$log.rt = log(d$Answer.train_rt)
  total = dim(d)[1]
  sd2 = 2*sd(d$log.rt)
  d = d[(d$log.rt > (mean(d$log.rt) - sd2)) & (d$log.rt < (mean(d$log.rt) + sd2)),]
  #print(paste("percent trimmed:" , round((total - dim(d)[1])/total,2)))
  
  # plot histogram
  sd = sd(d$log.rt)
  ggplot(d, aes(x=log.rt)) + 
    geom_histogram(fill = "black", alpha = .6) +
    geom_vline(aes(xintercept=mean(log.rt, na.rm=T)), 
               color="red", linetype="solid", size=.5) + 
    geom_vline(aes(xintercept=mean(log.rt, na.rm=T)-2*sd), 
               color="red", linetype="dashed", size=.5) +
    geom_vline(aes(xintercept=mean(log.rt, na.rm=T)+2*sd), 
               color="red", linetype="dashed", size=.5) +
  xlab("Log Study Time (ms)") +
  ylab("Frequency") +
  ggtitle('Histogram of study time') +
  theme(text = element_text(size=fs),
        plot.title=element_text(size=ts, face = "bold"),
        plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = 'black'))
  
  # get obj condition
  d = d[d$trial > 1,] # exclude anchors (not interesting because order not randomized)
  d$train_image2 <- as.character(matrix(lapply(str_split(d$Answer.train_image,"j"), function(x) {x[2]})))
  d$objCondition <- as.factor(as.character(matrix(lapply(str_split(d$train_image2,"-"), function(x) {x[1]}))))
  d$objItem <- as.character(matrix(lapply(str_split(d$train_image2,"-"), function(x) {x[2]})))
  d$objItem<- as.factor(as.numeric(gsub("[[:punct:]]", "", d$objItem)))
  d$obj <- paste("\"", d$objCondition, "-" ,d$objItem,"\"", sep = "" )
  d$train_image2 <- NULL

  # look at relationship between rt and memory peformance (no relationship)
  D = merge(d, da, by=c("workerid","objCondition","objItem"))
  # aggregate across participants
  D.agg <- ddply(D, .(workerid,Answer.test_answerEval), function (d) {mean(d$log.rt)})
  names(D.agg)[3] = "log.rt"
  
  # remove subjects that don't have both hits and misses for
  m = as.data.frame(table(D.agg$workerid))
  names(m)[1] = "workerid"
  goodIDs = m[which(m$Freq != 1),"workerid"]
  
  D.agg$good = ifelse(is.element(D.agg$workerid, goodIDs),1, 0)
  D.agg = D.agg[D.agg$good == 1,]

  ## stats: paired t-test
  # t.test(D.agg[D.agg$Answer.test_answerEval == "\"M\"","log.rt"],
        # D.agg[D.agg$Answer.test_answerEval == "\"H\"","log.rt"], paired = TRUE)

  # rt by item
  ms_all <- aggregate(log.rt  ~ obj, data=d, mean)
  ms_all$cih <- aggregate(log.rt ~ obj, data=d, ci.high)$log.rt
  ms_all$cil <- aggregate(log.rt ~ obj, data=d, ci.low)$log.rt
  
  # save RT by item
   #write.csv(ms_all, "data/rtNormsGeons_BYITEM.csv")
  #} 

rg_norms = read.csv("data/rtNormsGeons_BYITEM.csv")
```

Like for the complexity norms, study times were highly correlated with the number of geons in each object (_r_=.93, _p_<.01; see plot below, x-coordinates jittered to avoid over-plotting). Objects that contained more geons tended to be studied longer. 

```{r, echo = F, fig.width = 4, fig.height = 4}
## plot RT by condition
rg_norms$obj <- as.factor(as.numeric(gsub("[[:punct:]]", "", rg_norms$obj)))
rg_norms$obj_class = substr(rg_norms$obj, 1, 1)
rg_norms$obj_item = substr(rg_norms$obj, 2, 2)

# plot
ggplot(rg_norms, aes(y=log.rt, x=jitter(as.numeric(obj_class)))) +
  geom_pointrange(aes(ymax = log.rt+cih,ymin=log.rt-cil)) + 
  annotate("text", x=4.5, y=7.1, color = "red", size=rs,
           label=paste("r=",round(cor(as.numeric(rg_norms$obj_class), rg_norms$log.rt), 2))) +
  geom_smooth(method = "lm", color="blue", formula = y ~ x) +
  xlab("Object Condition") +
  ylab("Log study time (ms)") +
  ggtitle("Study time vs. # of geons") +
  theme(text = element_text(size=fs),
        plot.title=element_text(size=ts, face = "bold"),
        plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = 'black')) 
```

```{r, include = F}
## stats: relationship between reaction time norms and number of geons (BY ITEM)
cor.test(as.numeric(rg_norms$obj_class), rg_norms$log.rt)
```

Study times were also highly correlated with complexity norms. Objects that were rated as more complex tended to be studied longer.

```{r, echo = F, fig.width = 4, fig.height = 4, warning = F}
## plot RT by complexity rating
# merge
index <- match(cg_norms$obj, rg_norms$obj)
cg_norms$rt_meanRating <- rg_norms$log.rt[index]
cg_norms$rt_meanRating_cil <- rg_norms$cil[index]
cg_norms$rt_meanRating_cih <- rg_norms$cih[index]

# plot 
ggplot(cg_norms, aes(meanRating, rt_meanRating))+
  geom_point() + 
  geom_smooth(method = "lm", color="blue", formula = y ~ x) +
  geom_errorbarh(aes(xmin=meanRating-cil, 
                     xmax=meanRating+cih), size=0.2, colour="black") +
  geom_errorbar(aes(ymin=rt_meanRating-rt_meanRating_cil, 
                    ymax=rt_meanRating+rt_meanRating_cih), size=0.2, colour="black") +
  annotate("text", x=.9, y=7.1, 
           label=paste("r=",round(cor(cg_norms$rt_meanRating, cg_norms$meanRating, 
                                      use = "complete"), 2)), col="red", size = 6) +
  ylim(7,7.7) +
  xlim(0,1) +
  xlab("Complexity Norms") +
  ylab("Log study time (ms)") +
  ggtitle("Study time vs. Complexity norms") +
  theme(text = element_text(size=fs),
        plot.title=element_text(size=ts, face = "bold"),
        plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = 'black'))
```

```{r, include = F}
## stats: correlation between between complexity rating and RT
cor.test(cg_norms$rt_meanRating, cg_norms$meanRating)
```

Study times were did not predict memory performance. The study times for hits (correct "yes" responses; _M_ = 7.33 ) did not differ from misses (correct "no" responses; _M_ = 7.34; _t_(223) = .61, _p_=.54).


<a name="9"/>
<h3> Study 9: Real object study time task</h3> 

The task can be found <a href="http://langcog.stanford.edu/expts/MLL/refComplex/Experiment30/ref_complex_30.html" target="_blank"> here</a>.

We excluded subjects who performed below chance on the memory task (fewer than 30 correct out of 60). A response was counted as correct if it was a correct rejection or a hit. This excluded 6 subjects (1%). With these participants excluded, the mean correct was 84%.

Participants were also excluded based on study times. We transformed the time into log space, and excluded responses that were 2 standard deviations above or below the mean. This excluded 4% of responses. A histogram of study times is presented below. The solid line indicates the mean, and the dashed lines indicates two standard deviations above and below the mean.

```{r 9:objects_rt_norms, warning = F, echo = F, fig.width = 4, fig.height = 4, message = F}
## get data into long form (accuracy and RT df) and do exclusions
#if (processNorms){
  raw <- read.csv("data/RefComplex30.results",sep="\t",header=TRUE)
    
  # look at accuracy, and exclude those below chance
  # boxplot(raw$Answer.correct, main = "distribution before accuracy exclusions")
  raw = raw[raw$Answer.correct > 30, ]
  
  # accuracy dataframe
  n <- names(raw)
  cols = c( n[grepl("test",n)])
  mda <- melt(raw,id.vars=c("workerid"), measure.vars=cols,na.rm=TRUE)
  mda$trial <-as.numeric(matrix(lapply(str_split(mda$variable,"_"),function(x) {x[3]})))
  mda$var1 <- matrix(lapply(str_split(mda$variable,"_"),function(x) {x[1]}))
  mda$var2 <- matrix(lapply(str_split(mda$variable,"_"),function(x) {x[2]}))
  mda$var <- paste(mda$var1, mda$var2, sep = "_")
  mda$variable <- NULL; mda$var1 <- NULL; mda$var2 <- NULL
  
  mda$seq <- with(mda, ave(value, workerid, var, FUN = seq_along))
  da = dcast(workerid + seq + trial ~ var, data = mda, value.var = "value")
  da$seq <- NULL
  
  da=da[!is.na(da$trial),]
  
  da$Answer.test_answer = as.factor(da$Answer.test_answer)
  da$Answer.test_answerEval = as.factor(da$Answer.test_answerEval)
  da$Answer.test_image  = as.factor(da$Answer.test_image)
  
  # look at memory performance (correct: CR or hit)
  numCR = length(which(da$Answer.test_answerEval == "\"CR\""))
  numH = length(which(da$Answer.test_answerEval == "\"H\""))

  correct = (numCR + numH)/ dim(da)[1]
  #print(paste("percent correct:" , round(correct,2)))
  
  # RT dataframe
  n <- names(raw)
  cols = c(  n[grepl("train",n)], n["correct"])
  cols = cols[1:154]
  md <- melt(raw,id.vars=c("workerid"), measure.vars=cols,na.rm=TRUE)
  md$trial <-as.numeric(matrix(lapply(str_split(md$variable,"_"),function(x) {x[3]})))
  md$var1 <- matrix(lapply(str_split(md$variable,"_"),function(x) {x[1]}))
  md$var2 <- matrix(lapply(str_split(md$variable,"_"),function(x) {x[2]}))
  md$var <- paste(md$var1, md$var2, sep = "_")
  md$variable <- NULL; md$var1 <- NULL; md$var2 <- NULL
  
  md$seq <- with(md, ave(value, workerid, var, FUN = seq_along))
  d = dcast(workerid + seq + trial ~ var, data = md, value.var = "value")
  d$Answer.train_NA <- NULL; d$seq <- NULL
  
  d=d[!is.na(d$trial),]
  d=d[!is.na(d$Answer.train_image),]
  d=d[!is.na(d$Answer.train_rt),]
  d=d[d$Answer.train_image != "undefined",]
  d=d[d$Answer.train_image != "",]
  
  d$Answer.train_image = as.factor(d$Answer.train_image)
  d$Answer.train_rt = as.numeric(d$Answer.train_rt)
  
  #exclude outlier 2 standard deviations above and below mean (in log space)
  d$log.rt = log(d$Answer.train_rt)
  total = dim(d)[1]
  sd2 = 2*sd(d$log.rt)
  d = d[(d$log.rt > (mean(d$log.rt) - sd2)) & (d$log.rt < (mean(d$log.rt) + sd2)),]
  #print(paste("percent trimmed:" , round((total - dim(d)[1])/total,2)))
  
  # plot histogram
  sd = sd(d$log.rt)
  ggplot(d, aes(x=log.rt)) + 
    geom_histogram(fill = "black", alpha = .6) +
    geom_vline(aes(xintercept=mean(log.rt, na.rm=T)), 
               color="red", linetype="solid", size=.5) + 
    geom_vline(aes(xintercept=mean(log.rt, na.rm=T)-2*sd), 
               color="red", linetype="dashed", size=.5) +
    geom_vline(aes(xintercept=mean(log.rt, na.rm=T)+2*sd), 
               color="red", linetype="dashed", size=.5) +
  xlab("Log Study Time (ms)") +
  ylab("Frequency") +
  ggtitle('Histogram of study time') +
  theme(text = element_text(size=fs),
        plot.title=element_text(size=ts, face = "bold"),
        plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = 'black'))
  
  # look at relationship between rt and memory peformance (no relationship)
  D = merge(d, da, by.x=c("workerid","Answer.train_image"), by.y = c("workerid","Answer.test_image"))
  D = D[D$Answer.train_image != 61,] # get rid of ball and motherboard
  D = D[D$Answer.train_image != 62,] 
  D_ms = ddply(D ,.(Answer.test_answerEval), function (d) {mean(d$log.rt)})
  # aggregate across participants
  D.agg <- ddply(D, .(workerid,Answer.test_answerEval), function (d) {mean(d$log.rt)})
  names(D.agg)[3] = "log.rt"
  
  # remove subjects that don't have both hits and misses for
  m = as.data.frame(table(D.agg$workerid))
  names(m)[1] = "workerid"
  goodIDs = m[which(m$Freq != 1),"workerid"]
  
  D.agg$good = ifelse(is.element(D.agg$workerid, goodIDs), 1, 0)
  D.agg = D.agg[D.agg$good == 1,]
  means = ddply(D.agg, .(Answer.test_answerEval), function (d) {mean(d$log.rt)})

  ## stats: paired t-test of RTs between misses and hits  
  # t.test(D.agg[D.agg$Answer.test_answerEval == "\"M\"","log.rt"],
       #D.agg[D.agg$Answer.test_answerEval == "\"H\"","log.rt"], paired = TRUE)

  # aggregate
  ms <- aggregate(log.rt  ~ Answer.train_image, data=d, mean)
  ms$rt_cil <- aggregate(log.rt ~ Answer.train_image, data=d, ci.low)$log.rt 
  ms$rt_cih <- aggregate(log.rt ~ Answer.train_image, data=d, ci.high)$log.rt 
  
  ms = ms[ms$Answer.train_image != 61,] # get rid of ball and motherboard
  ms = ms[ms$Answer.train_image != 62,] 

  #write.csv(ms,"data/rtNormsObjs_BYITEM.csv")
  #}

rto_norms = read.csv("data/rtNormsObjs_BYITEM.csv")
```

The plot below shows the correlation between study time and explicit complexity norms for each object. Like for the geons, objects that were rated as more complex were studied longer.

```{r, echo = F, fig.width = 4, fig.height = 4, warning = F}
## plot complexity ratings vs RTs
# merge
index <- match(co_norms$ratingNum, rto_norms$Answer.train_image)
co_norms$log.rt <- rto_norms$log.rt[index]
co_norms$rt_meanRating_cil <- rto_norms$rt_cil[index]
co_norms$rt_meanRating_cih <- rto_norms$rt_cih[index]

# plot
ggplot(co_norms, aes(meanRating, log.rt))+
  geom_point() + 
  geom_smooth(method = "lm", color="blue", formula = y ~ x) +
  geom_errorbarh(aes(xmin=meanRating-cil, 
                     xmax=meanRating+cih), size=0.2, colour="black") +
  geom_errorbar(aes(ymin=log.rt-rt_meanRating_cil, 
                    ymax=log.rt+rt_meanRating_cih), size=0.2, colour="black") +
  annotate("text", x=.9, y=7.1, 
           label=paste("r=",round(cor(co_norms$meanRating, co_norms$log.rt,
                                      use = "complete"), 2)), col="red", size = 6) +
  ylim(7,7.7) +
  xlim(0,1) +
  xlab("Complexity Norms") +
  ylab("Log study time (ms)") +
  ggtitle("Study times vs. Complexity norms") +
  theme(text = element_text(size=fs),
        plot.title=element_text(size=ts, face = "bold"),
        plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = 'black'))
```

```{r include = F}
## stats: correlation between between complexity rating and RT
cor.test(co_norms$meanRating, co_norms$log.rt)
```
For the real objects, study times predicted memory performance. Study times for hits (correct yes responses; _M_ = 7.24) were greater than for misses (correct no responses; _M_ = 7.11; _t_(393) = 9.74, _p_<.0001).

<a name="10"/>
<h3> Study 10: English complexity norms</h3> 

The task can be found <a href="http://langcog.stanford.edu/expts/MLL/refComplex/Experiment26/ref_complex_26.html" target="_blank"> here</a>.

TO DO

```{r 10:english_norms, warning= F, include = F}
## preprocess and merge in stuff
if (processNorms) {
  d1 <- read.csv("data/RefComplex26.results1",sep="\t",header=TRUE)
  d2 <- read.csv("data/RefComplex26.results",sep="\t",header=TRUE)
  d <- rbind(d1,d2)
  
  # take out people who missed check question
  d = d[d$Answer.value_18 == 7, ] 
  
  # melt into word and values into two data frames, then rejoin based on workerid+trial id
  # (tricky because variable is string and number)
  n <- names(d)
  colsV =  n[grepl("value",n)] 
  colsW =  n[grepl("word_",n)] 
  mdW <- melt(d,id.vars=c("workerid"), measure.vars=colsW,na.rm=TRUE)
  mdW$trial <- as.numeric(matrix(lapply(str_split(mdW$variable,"_"),function(x) {x[2]})))
  mdW$index <- paste(mdW$workerid, mdW$trial, sep="_")
  mdV <- melt(d,id.vars=c("workerid"), measure.vars=colsV,na.rm=TRUE)
  mdV$trial <- as.numeric(matrix(lapply(str_split(mdV$variable,"_"),function(x) {x[2]})))
  mdV$index <- paste(mdV$workerid, mdV$trial, sep="_")
  
  # merge together
  index <- match(mdW$index, mdV$index)
  mdW$complexity <- mdV$value[index]
 
  # delete variables
  mdW$index <- NULL; mdW$variable <- NULL; names(mdW)[2] <- "word"; md <- mdW
  
  # remove quotes from words
  md$word= gsub(" ", "", gsub("[[:punct:]]", "", md$word))
  md$word = as.factor(md$word)
  
  # remove non-words
  md = md[md$word != "ball",] # anchor
  md = md[md$word != "motherboard",] # anchor
  md = md[md$word != "43",] # take out check question
  md = md[md$word != "peso",] # take out bad word
  
  # add length in characters
  md$nchars = nchar(as.character(md$word))
  
  # merge in word info
  # -- add mrc data --
  mrc = read.csv("data/MRC_corpus.csv")
  mrc = mrc[mrc$mrc.syl != "NA",]
  index <- match(md$word, mrc$word)
  md$mrc.fam <- mrc$mrc.fam[index]
  md$mrc.conc <- mrc$mrc.conc[index]
  md$mrc.imag <- mrc$mrc.imag[index]
  md$mrc.syl <- mrc$mrc.syl[index]
 
  # -- add phonemes from MRC --
  mrc = mrc[mrc$mrc.wtype != " ",]
  index <- match(md$word, mrc$word)
  md$mrc.phon <- mrc$mrc.phon[index]

  # -- add class --
  class = read.csv("data/english_class_codes.csv")
  index <- match(md$word, class$ENGLISH)
  md$class <- class$class_MLL[index]
  
  # -- add morphemes --
  morph = read.csv("data/numMorph_celex2.csv")
  index <- match(md$word, morph$ENGLISH)
  md$clx.morph <- morph$clx.numMorph[index]
  
  # -- add frequency --
  freqs = read.table("data/SUBTLEXusDataBase.txt",header=TRUE)
  index <- match(md$word, freqs$Word)
  md$subt.log.freq <- freqs$Lg10WF[index]
  
  # -- add brysbaert concreteness --
  b <- read.csv("data/brysbaert_corpus.csv",header=TRUE)
  b <- b[b$Word != "",] # get rid of empty rows
  b <- b[b$Bigram == 0,]# get rid of two word lemmas
  index <- match(md$word, b$Word)
  md$b.conc <- b$Conc.M[index]
  
  write.csv(md, "data/englishComplexityNorms.csv")
 }

eng = read.csv("data/englishComplexityNorms.csv")
```

```{r, echo = F, fig.width = 4, fig.height = 4}
## plot complexity norms vs word length
# aggregate across words of the same length
ms2 <- aggregate(complexity ~ nchars, data=eng, mean)
ms2$cih <- aggregate(complexity ~ nchars,  data=eng, ci.high)$complexity
ms2$cil <- aggregate(complexity ~ nchars,  data=eng, ci.low)$complexity

# plot 
ggplot(ms2, aes(complexity, nchars)) +
  geom_point() + 
  geom_smooth(method = "lm", color="blue", formula = y ~ x) +
  geom_errorbarh(aes(xmax=complexity+cih, xmin=complexity-cil), size=.2, colour="black") +
  annotate("text", x=6, y=3, size = rs, 
           label=paste("r=",round(cor(eng$complexity, eng$nchars), 2)), col="red") +
  scale_y_continuous(limits = c(0, 15), breaks = 1:14, labels = 1:14) +
  scale_x_continuous(limits = c(0, 7), breaks = 1:7, labels = 1:7)  +
  xlab('Complexity Rating') +
  ylab('Word Length (chars.)') +
  ggtitle("# of characters vs. complexity rating") +
  theme(text = element_text(size=fs),
       plot.title=element_text(size=ts, face = "bold"),
       plot.background = element_blank(),
       panel.grid.major = element_blank(),
       panel.grid.minor = element_blank(),
       panel.border = element_blank(),
       axis.line = element_line(color = 'black'))
```

```{r, include = F}
## aggregate across participants
# have to do this separately for each length variable because drops words for which there are NAs *for any of the length vars*
ms.syl <- aggregate(complexity ~ word + mrc.syl, data=eng, mean)
ms.phon <- aggregate(complexity ~ word + mrc.phon, data=eng, mean)
ms.morph <- aggregate(complexity ~ word + clx.morph, data=eng, mean)
ms.chars <- aggregate(complexity ~ word + nchars +subt.log.freq + mrc.conc, data=eng, mean)
```

```{r, include = F}
## stats: correlations between length and complexity in English
# correlation between syllables and complexity
cor.test(ms.syl$complexity, ms.syl$mrc.syl)

# correlation between phonemes and complexity
cor.test(ms.phon$complexity, ms.phon$mrc.phon)

# correlation between morphomes and complexity
cor.test(ms.morph$complexity, ms.morph$clx.morph)

# mono-morphemic
ms_mono.syl = aggregate(complexity ~ word + mrc.syl, data=eng[eng$clx.morph == 1,], mean)
ms_mono.phon = aggregate(complexity ~ word + mrc.phon, data=eng[eng$clx.morph == 1,], mean)

# correlation between syllables and complexity [mono-morphemic only]
cor.test(ms_mono.syl$complexity, ms_mono.syl$mrc.syl)

# correlation between phonemes and complexity [mono-morphemic only]
cor.test(ms_mono.phon$complexity, ms_mono.phon$mrc.phon)

# open class
ms_open.syl = aggregate(complexity ~ word + mrc.syl, data=eng[eng$class != 0,], mean)
ms_open.phon = aggregate(complexity ~ word + mrc.phon, data=eng[eng$class != 0,], mean)
ms_open.morph = aggregate(complexity ~ word + clx.morph, data=eng[eng$class != 0,], mean)

# correlation between syllables and complexity [open class only]
cor.test(ms_open.syl$complexity, ms_open.syl$mrc.syl)

# correlation between phonemes and complexity [open class only]
cor.test(ms_open.phon$complexity, ms_open.phon$mrc.phon)

# correlation between morphemes and complexity [open class only]
cor.test(ms_open.morph$complexity, ms_open.morph$clx.morph)

# paritialing out frequency
pcor.test(ms.chars$complexity,ms.chars$nchars,ms.chars$subt.log.freq)

# concreteness split
ms.chars <- aggregate(complexity ~ word + nchars + mrc.conc, data=eng, mean) # do this again without freq because drops missing words
ms.chars$conc.split = ifelse(ms.chars$mrc.conc < median(ms.chars$mrc.conc), 1, 2)
ms.chars$conc.split=as.factor(ms.chars$conc.split)
ms.charsL = ms.chars[ms.chars$conc.split == 1,]
cor.test(ms.charsL$nchars, ms.charsL$complexity)

ms.chars <- aggregate(complexity ~ word + nchars + mrc.conc + subt.log.freq, data=eng, mean) # do this again without freq because drops missing words
ms.chars$conc.split = ifelse(ms.chars$mrc.conc < median(ms.chars$mrc.conc), 1, 2)
ms.charsL = ms.chars[ms.chars$conc.split == 1,]
pcor.test(ms.charsL$complexity,ms.charsL$nchars,ms.charsL$subt.log.freq)
```


```{r, include = F}
## are the correlations between length and complexity reliable, controlling for everything? [yes]
# all
m_syl = lm(mrc.syl ~ complexity + mrc.fam + mrc.imag + b.conc + subt.log.freq, eng)
m_phon = lm(mrc.phon ~ complexity + mrc.fam + mrc.imag + b.conc + subt.log.freq, eng)
m_morph = lm(clx.morph ~ complexity + mrc.fam + mrc.imag + b.conc + subt.log.freq, eng)
summary(m_syl)
summary(m_phon)
summary(m_morph)

# mono-morphemic
m_syl_m = lm(mrc.syl ~ complexity + mrc.fam + mrc.imag + b.conc + subt.log.freq, eng[eng$clx.morph == 1,])
m_phon_m = lm(mrc.phon ~ complexity + mrc.fam + mrc.imag + b.conc + subt.log.freq, eng[eng$clx.morph == 1,])
summary(m_syl_m)
summary(m_phon_m)

# open class
m_syl_o = lm(mrc.syl ~ complexity + mrc.fam + mrc.imag + b.conc + subt.log.freq, eng[eng$class != 0,])
m_phon_o = lm(mrc.phon ~ complexity + mrc.fam + mrc.imag + b.conc + subt.log.freq, eng[eng$class != 0,])
m_morph_o = lm(clx.morph ~ complexity + mrc.fam + mrc.imag + b.conc + subt.log.freq, eng[eng$class != 0,])
summary(m_syl_o)
summary(m_phon_o)
summary(m_morph_o)
```

```{r, include = F}
## create english word data frame with relevant variables for xling analysis
xling.eng <- aggregate(complexity ~ word + class, data=eng, mean)

# -- add morphemes --
morph = read.csv("data/numMorph_celex2.csv")
index <- match(xling.eng$word, morph$ENGLISH)
xling.eng$clx.morph <- morph$clx.numMorph[index]

# -- add frequency --
freqs = read.table("data/SUBTLEXusDataBase.txt",header=TRUE)
index <- match(xling.eng$word, freqs$Word)
xling.eng$subt.log.freq <- freqs$Lg10WF[index]

xling.eng$class = as.factor(xling.eng$class)
xling.eng$clx.morph = as.factor(xling.eng$clx.morph)
```

<a name="11"/>
<h3> Study 11: Cross-linguistic analysis</h3> 

TO DO

<a name="12"/>
<h3> Study 12: Simultaneous frequency task</h3> 

The task can be found <a href="http://langcog.stanford.edu/expts/MLL/refComplex/Experiment17_2/ref_complex_17_2.html" target="_blank"> here</a>.

Plotted below is the proportion of low frequency object selections as a function of language condition (long vs. short). Selections between the two conditions did not differ.

```{r 12: simul_freq, include = F}
# read in data
d_simultaneous <- read.csv("data/RefComplex6_all.results", header=TRUE)
```

```{r, echo = F, fig.width = 4, fig.height = 4}
## plot proportion low freq by language condition
# get proportions by condition
ms_simult = ddply(d_simultaneous ,.(Answer.lang_condition), function (d, dv) {p.fc(d, "\"low_freq\"")})
ms_simult$Answer.lang_condition = c("long", "short")

# plot
qplot(Answer.lang_condition, p_complex, fill = Answer.lang_condition,
      ylim = c(0,1), position=position_dodge(),
      data = ms_simult, geom="bar", 
      ylab = "Prop. selection low freq. object", xlab="Language condition", 
      main = "Low freq. selections vs. language", stat="identity")  +
  geom_linerange(aes(ymin=ciwl,ymax=ciul), position=position_dodge(.9)) +
  geom_abline(intercept = .5, slope = 0, linetype = 2) +
  theme(text = element_text(size=fs),
        plot.title=element_text(size=ts, face = "bold"),
        plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = 'black'),
        legend.position="none")
```

```{r, include = F}
## stats: proportion low freq. in long vs. short condition
chisq.test(table(d_simultaneous$responseValue, d_simultaneous$Answer.lang_condition))
```

<a name="13"/>
<h3> Study 13: Sequential frequency task</h3> 

The task can be found <a href="http://langcog.stanford.edu/expts/MLL/refComplex/Experiment16/ref_complex_16.html" target="_blank"> here</a>.

Plotted below is the proportion of low frequency object selections as a function of language condition (long vs. short). Selections between the two conditions did not differ.

```{r, include = F}
## read in data and pre-process
d_sequential <- read.csv("data/RefComplex16.results",sep="\t",header=TRUE)

d_sequential <- d_sequential[d_sequential$Answer.frequency_condition == '\"uneven\"', ]
levels(d_sequential$Answer.crit_selection) = c("\"high_freq\"", "\"low_freq\"")
names(d_sequential)[35] = "responseValue"
```

```{r 13:sequential_freq, echo = F, fig.width = 4, fig.height = 4}
## plot proportion low freq by language condition
# get proportions by condition
ms_seq = ddply(d_sequential,.(Answer.lang_condition), function (d, dv) {p.fc(d, "\"low_freq\"")})
ms_seq$Answer.lang_condition = c("long", "short")

# plot
qplot(Answer.lang_condition,p_complex, fill = Answer.lang_condition,
      ylim=c(0,1), position=position_dodge(),
      data=ms_seq,geom="bar", ylab="Prop. selection low freq. object", 
      main = "Low freq. selections vs. language",
      xlab="Language condition", stat="identity") +
  geom_linerange(aes(ymin=ciwl,ymax=ciul), position=position_dodge(.9)) +
  geom_abline(intercept = .5, slope = 0, linetype = 2) +
  theme(text = element_text(size=fs),
        plot.title=element_text(size=ts, face = "bold"),
        plot.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.border = element_blank(),
        axis.line = element_line(color = 'black'),
        legend.position="none")
```

```{r, include = F}
## stats: proportion low freq. in long vs. short condition
chisq.test(table(d_sequential$responseValue, d_sequential$Answer.lang_condition))
```